|

### 2021-09-28 _main.Rmd
End of day:

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy in Population Health Sciences
degreedate: TBC
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.md"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: For Brendan McHugh
covidStatment: |
  `r paste(readLines("front-and-back-matter/_covidStatment.Rmd"), collapse = '\n  ')`
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations

#######################
## bibliography path ##
#######################
bibliography: 
    - "bibliography/references.bib"
csl: bibliography/nature.csl
bibliography-heading-in-pdf: Bibliography

@@

Allowing users of `medrxivr` to access a maintained snapshot removes any dependency on the API, which can become unavailable during peak usage times. The relationship between the two methods for accessing the data contained in the medRxiv database is summarized in Figure \@ref(fig:medrxivr-data-sources).

(ref:medrxivr-data-sources-cap) **Overview of `medrxivr` data sources:** Users can either access the API directly via `mx_api_content()`, or can import a maintained snapshot of the database, taken each morning at 6am, via the `mx_snapshot()` function. Note: due to the size of bioRxiv, only a maintained snapshot of the medRxiv repository is available via `mx_snapshot()`.

@@

### Refining your search

An important argument of the `mx_search()` is `report`, which outputs a structured table with each search strategy presented on an individual line and the number of records associated with this strategy. __[CITATION NEEDED - PRISMA-S]__

```{r, eval = FALSE}
results  <- mx_search(mx_data,
                      myquery, # Combined with OR
                      report = TRUE)
```

@@

The package was been onboarded into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web".[@boettiger2015] As part of this process, following rigorous peer-review, an associated article introducing the tool was published by the Journal of Open Source Software.[@mcguinness2020a] The entire review discussion is publicly available and can be viewed online. [^chp-systematic-review-tools-4] The tool has also been well received by the open-source community, demonstrated by the engagement of other developers in contributing to important new functionality and suggesting bug-fixes.

[^chp-systematic-review-tools-4]: <https://github.com/ropensci/software-review/issues/380>

@@

### Limitations of `medrxivr`

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of this thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, meaning some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medrxiv/bioRxiv is substantially different from the true grey literature (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979] It is likely too early (and likely too methodologically difficult) to tell whether the increased popularity and acceptance of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues.

@@

<!-- -   TODO Examine how many times a primary study was included in a review -->
<!-- -   TODO Look at whether searching preprints made any difference to my results -->
<!-- -  TODO Could apply ROBIS to previous reviews to see how they stack-up? (This may be done as part of Georgia's summer project) -->

<!-- TODO Extra studies to include: -->

<!-- * Smeet et al. reference in the CPRD analysis -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->
<!-- * Statins and the risk of dementia Jick -->

@@

The importance of

I found that

@@

The aim of this chapter is to systematically review all available literature on the association between blood levels of total cholesterol and it's constituent parts (HDL-c,LDL-c and triglycerides) on the subsequent risk of dementia.

Based on the review of existing literature, no previous evidence synthesis exercise attempted to examine all available literature simultaneously.

Only by looking across all available evidence can the true, given the often limited use of one type of evidence (RCTs are hard to do, obs studies are biased).

Literature con

Systematic reviews aim to transparently and reproducibly identift

<!-- TODO Need to seriously improve this -->

@@

This systematic review was performed in conjunction with a team of secondary reviewer, all of whom . These included Alexandra MacAleenana, Athena Sheppard, Matthew Lee and Lena Schmidt.

@@

### Search strategy and and study selection

I systematically search electronic bibliographic databases to identify potentially relevant records. The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy will include terms related to lipids, lipid modifying treatments, and dementia and its sub-types, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full draft search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection (for the specific databases available thi via the University). We will also search clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials. As the contents of the Web of Science Core COllection can vary by institution, the specific database searched via this platform are listed in Appendix \@ref(appendix-wos-databases).

The abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) were searched by hand. <!-- TODO how were these searched? Date and other range -->

 Grey literature will also be searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses were accessed using the Open Access Theses and Dissertations portal. We will also search bioRxiv and medRxiv, preprint repositories using a tool built as part of this thesis, to identify potentially relevant studies. Finally, the reference lists of included studies were searched by hand while studies citing included studies will be examined using Google Scholar (forward and reverse citation searching).

<!-- TODO Look at citationchaser for this, but have a cut-off publication date of July 2019 - better than searching by hand/Google Scholar -->

<!-- Include table with overview of search strategy here: summarise topics and how they were combined -->

#### Study selection

Records were imported into Endnote and deduplicated using the method outlined in Bramer et al. (2016).[@bramer2016] In summary, this method uses multiple stages to identify potential duplicates, beginning with automatic deletion of records matching on multiple fields ("Author" + "Year" + "Title" + "Journal"), followed by manual review of less similar articles (e.g. those matched based on the "Title" field alone).

Screening (both title/abstract and full-text) was performed using a combination of Endnote, a citation management tool,[@hupe2019] and Rayyan, a web-based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records was performed primarily by the primary author, with a random selection of excluded records being screened in duplicate to ensure consistency with the inclusion criteria.

Full-text screening will also be completed in full by the primary author. A second reviewer screened a random sample (approx 10%) of included and excluded records, in addition to any records identified by the first reviewer as being difficult to assess against the inclusion criteria. Reasons for exclusion at this stage were recorded. Disagreements occurring during either stage of the screening process were resolved through discussion with a senior colleague. A PRIMSA flow diagram will be produced to document how records moved through the review.[@page2021]

@@

We will seek studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia/MCI. Eligible study designs include randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants were free (or assumed to be free) of dementia/MCI at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits will be placed on the sample size of included studies.

Eligible studies defined dementia according to recognised criteria, for example the National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN), International Classification of Diseases (ICD), or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria. <!-- TODO do i actually want to include MCI? -->
 For MCI, eligible studies are those that attempted state a definition for diagnoses of MCI (e.g. an adapted version of the Petersen criteria [@petersen1999]) and create ordinal groups of patients (e.g. no dementia or dementia/MCI/dementia) based on this definition.

No limitations will be imposed on publication status, publication date, venue or language, although sufficiently detailed reports of the studies to be able to examine their methods were required for inclusion.

@@

Case-control studies, cross-sectional studies, qualitative studies, case reports/series and narrative reviews will be excluded. Studies which present no evidence of attempting to exclude prevalent cases from their analyses will also be excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/MCI/dementia) will be excluded. Conference abstracts with no corresponding full-text publication will be examined, and we will contact authors to obtain information on the study's status. Studies that are reported in insufficient detail (e.g. only in conference abstracts, new, letters, editorials and opinion) will be excluded. Previous systematic reviews will not be eligible, but their reference lists will be screened to identify any potentially relevant articles. Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a "multi-domain intervention" where the lipid-regulating agent is included in each arms (e.g. for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e. below the age of 50) will be excluded.

Excluded studies performing autopsy unless it was done under accepted criteria Exclude studies using a dietary intervention, for example omega-3 fatty acid enriched diet, as it hard to disentangle the effect of other elements contained within te diet, vs simple tablet based supplements of

<!-- TODO Expand on rationale for exclusion of cross-sectional studies. Dementia is a disease severly at risk of differential recall, leading to misclassification bias. By using prospective studies, the potential for this bias is reduced. -->

#### Assessment of inter-rater reliability

Inter- and intra-rater reliability was assessed for a 10% sub-sample of records at the title and abstract screening stage. Intra-rater reliability involved a single reviewer applying the inclusion criteria to the same set of records while blinded to their previous decisions, while inter-rater reliability involved two reviewers independently screening the same set of records.

If this approach demonstrates a significant level of erroneous exclusion by the primary author a larger proportion will be dual-screened.

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure of inter-rater reliability was chosen over other methods of assessing inter-rater reliability such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review, this is an important characteristic.

@@

@@

Harmonization of cholesterol measures across studies was performed, as different studies used different methods to quantify exposure, including comparing differing risks in the highest vs lowest quartiles of a lipid, using a binary classification of patients into a hypercholesterolaemia or not, categorizing lipid levels into high, middle, and low groups according to study-defined criteria, and simply treating the exposure as a continuous variable.

Following guidance for a dose response meta-analysis, I picked midpoint for all studies where bands were provided (extremely common in the case of lipids)

This analysis was unusual, as multiple methods of integration and analysis were required with data spanning many forms.

A piloted data extraction for was used. I extracted the data in the first instance and subset was checked by a colleague for accuracy.

<!-- TODO Cross reference with the data extraction issues here - i.e. not providing qratile cut-off values, not poviding numbers per group, etc. -->

Where necessary, lipid levels reported in _mmol/L_ were converted in _mg/dL_ using the following formula:

@@

#### Studification

Multiple publications resulting from the analysis of the same data were included and grouped ().

In the advent of multiple papers reporting results on the same cohort, but say, at different time points, I used a process of studifiction to build out the most comprehensive accounts of the studies and results from as many published articles were applicable.

This was particularly relevant to preprints and published papers reporting the same study, which were not considered duplicate records. Instead, they were considered as different reports of the same study. This is due to the potential for the published version to offer some information that the preprint did not, and vice versa.

@@

Where data points required either for the analysis or risk-of-bias assessment but were not reported, the primary authors of the study were contacted. This approach was taken due to the potentially large impact of following up with authors on the results of the review,[@reynders2019]

This was particular important for the dose response meta-analysis, we

Use the example of getting the cut-off points from the

<!-- TODO move to results, and keep just a short bit here -->

#### Converting between different

Include formulae and informed assumptions

Not interchangeable in the context of systematic review. If outcome is rare, odds and risk ratios approximate each other, but the hazard ratio is measuring something completely different, by taking into account time-to-event in each treatment group.

Several existing reviews do not distinguish between the two types of effect measures and include all existing studies in a single meta-analysis to produce an overall effect.

The likely effect of this is that the overall effect measures are biased towards . . . . ?

Can you predict the direction of effect from a

Best practice methods for dealing with disparate effect measures in included studies is to not perform a meta-analysis

<!-- See the third row of the table here; https://training.cochrane.org/handbook/current/chapter-12#section-12-1 -->

Some evidence of manipulation of effect estimates in previous reviews,(e.g. Chou, Sci Reports - at least one study disagrees with) but not documented in review text.

WHile the analysis of change scores is feasible, as they were explicityl

### Risk of bias assessment {#risk-of-bias}

A key use of the review presented in this chapter is to identify different sources of evidence at risk of a diverse range of biases, and to contrast and compare findings across them (see Section \@ref(intro-triangulation) and Section \@ref(triangulation-overview) for an overview of triangulation). To enable this triangulation exercise, a detailed and structured risk of bias assessment formed an important part of this review.

@@

This move was prompted by a unclear definition of "methodological quality" which could include facets such as, and challenges in the comparison of results from different tools.

As part of this, the community also moved from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order.

Additionally, results should be assessed at the result (defined as a a specific outcome at a specific timepoint) rather than the study level. For example, a study may report on the efficacy of an intervention at 6 months and two years follow-up. missing outcome data that is not an issues at 6 months may introduce bias at 2 year follow-up. In this case, assigning a bias judgement to the study as a whole hides this differential biases for each result.

@@

In addition, the tool also aim to capture the potential direction of bias for each given domain. An optional question at the end of each domain, e.g. for Domain 1 of the RoB2 tool: "What is the predicted direction of bias arising from the randomization process?", possible responses included

* Not applicable 
* Favours experimental 
* Favours comparator
* Towards null 
* Away from null 
* Unpredictable

@@

Studies were not combined across different study designs (i.e. RCTs were not combined in a meta-analysis with results from observational studies).

>>>>

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy
degreedate: 2022
word-count: |
  `r read.csv("data/words/words.csv") %>% dplyr::pull(1) %>% dplyr::last()`
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.md"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: For Brendan McHugh
covidStatment: |
  `r paste(readLines("front-and-back-matter/_covidStatment.Rmd"), collapse = '\n  ')`
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations

#######################
## bibliography path ##
#######################
bibliography: [bibliography/references.bib, bibliography/packages.bib]
csl: bibliography/nature.csl
bibliography-heading-in-pdf: Bibliography

@@

This snapshot is created each morning at 6am using a process known as "git-scraping",[@zotero-15031] whereby the entire database is downloaded using the `mx_api_content()` function and saved as a comma separated value (CSV) file to an online server (Figure \@ref(fig:medrxivr-data-sources)). Calling `mx_snapshot()` imports this CSV into R, and has the advantage of both faster loading of the data into R (as it is imported as a single file and does not require cycling through the output of the API) and an absence of any reliance on the API.

The one limitation of this approach is that the snapshot (by its nature) will not contain details of records added to the database since it was taken. However, given that the number of records added each day is relatively low, this should pose minor issues.

(ref:medrxivr-data-sources-cap) **Overview of `medrxivr` data sources** - Users can either access the API directly via `mx_api_content()`, or can import a maintained snapshot of the database, taken each morning at 6am, via the `mx_snapshot()` function. Note: due to the size of bioRxiv, only a maintained snapshot of the medRxiv repository is available via `mx_snapshot()`.

@@

### Refining a search

An important argument of the `mx_search()` is `report`, which outputs a structured table with each search strategy presented on an individual line and the number of records associated with this strategy.[@rethlefsen2021prisma]

```{r, eval = FALSE}
results  <- mx_search(mx_data,
                      myquery,
                      report = TRUE)
```

@@

The package has been accepted into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web".[@boettiger2015] As part of this process, following rigorous peer-review, an associated article introducing the tool was published by the Journal of Open Source Software.[@mcguinness2020a] The entire review discussion is publicly available and can be viewed online.[@zotero-15016] The tool has also been well received by the open-source community, demonstrated by the engagement of other developers in contributing to important new functionality and suggesting bug-fixes.

@@

### Use case

In addition to being used to search systematically search health-related preprint servers, as illustrated in the systematic review presented in Chapter \@ref(sys-rev-heading), `medrxivr` has other uses. For example, I led a descriptive analysis of the change in data availability statements between preprinted and published versions of the same manuscript, stratified by journal data sharing policy access, underpinned by preprint meta-data provided by `medrixvr`.

By comparing the preprinted and published versions of the data availability statement, I could examine the same manuscript (same content, authors and funders) under two differnt publication policies, and examine whether stricter policies which require data sharing as a condition of publication actually result in increased data availability. We found some evidence that data availability statements more frequently described open data on publication when the journal mandated data sharing compared to when the journal did not mandate data sharing. This study has since been published in PLOS One, and a copy is included in Appendix \@ref(published-papers). Author contributions are discussed in the

In short, this use case illustrates that easy access to medRxiv/bioRxiv metadata has applications beyond systematic searching of preprints as a part of evidence synthesis exercises.

### Limitations of `medrxivr` {#medrxivr-limitations}

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, meaning some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

There is also the potential that the cross-section of literature posted on medrxiv/bioRxiv is substantially different those suffering from publication bias (studies or analyses that are not published for a range of reasons including results that are not deemed "novel" or are not statistically significant).[@song2010] This is because simply lowering the barriers to publication may well encourage authors to published "null" results, but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[@rosenthal1979]

<!-- TODO Clean-up this section -->

<!-- REVIEW I don’t think anyone would assume that preprints would help much here… I’d have thought the main argument is getting hold of evidence in a more timely manner.-->

```{r, include=FALSE}
med_res <-
  # Use snapshot from July 2021
  medrxivr::mx_search(medrxivr::mx_snapshot("ccedfb8a44304b9fba4e3ba518a8ce4ed2294770"),
                      "*",
                      from_date = "2019-07-01",
                      to_date = "2019-08-01") %>%
mutate(pub_ind = ifelse(published == "NA", 0, 1)) %>%
  group_by(pub_ind) %>%
  count()

published <- med_res$n[2]
publication_n <- sum(med_res$n)
publish <- paste0(num_to_text(published,T)," (", round(published/publication_n*100,1),"%)")
```

<!-- TODO Need to finish this bit and move it to the use-case section! -->

Using `medrxivr` an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r publish` of the `r publication_n` records posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting,[@abdill2019b] indicating that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

<!-- TODO replicate this analysis -->

It is likely too early (and likely too methodologically difficult) to tell whether the increased popularity and acceptance of preprint repositories will have any effect of the availability of research that was not considered "publishable" at other venues.

@@

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->
<!-- * Statins and the risk of dementia Jick -->

@@

Systematic reviews are a type of research that aim to use all existing evidence to provide the best answer to an important research question. They do this by finding and combining the results from many related primary research studies. Reviews involve multiple steps including: searching of existing studies; assessment of the studies against predefined inclusion criteria; collection of data from each study; assessment of each study's methods.

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels, and dementia.

There were XXX primary studies <!-- TODO check this -->that contained information on this relationship. I found that statins reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipids were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

The use of the results of this review in subsequent chapters is dicussed.

<!-- TODO Search for MCI -->

<!-- TODO text -->

In this chapter, I describe a comprehensive systematic review of the relationship between blood lipid levels, and treatments that modify them, and the subsequent risk of dementia and related outcomes. This analysis sought to address two specific aims.

Firstly, as discussed in the Introduction to this thesis (Section \@ref(evidence-association)), several diverse forms of evidence on the relationship of lipids and dementia exist. These include randomised controlled trials, observational studies of different analytical design, and Mendelian randomisation studies. However, based on a scoping review of existing literature, no previous evidence synthesis exercise has attempted to examine the association of lipids/statins with dementia outcomes across these distinct evidence types. Collating these diverse evidence sources is important, as if the observed association between lipids and dementia is constant across them, it increases our confidence in the association. As such, the primary aim of this analysis was to systematically review all available literature, regardless of study design.

Secondly, I explicitly sought to include health-related preprint servers as a potential evidence source in this review, as they are infrequently considered by evidence synthesists but may contain important unpublished studies. As a sensitivity analysis to the systematic review presented in this chapter, I sought to quantify the additional evidential value of including preprints. This inclusion of preprint serves makes use of the preprint search tool presented in Chapter \@ref(sys-rev-tools-heading).

@@

@@

In line with best-practice guidance, secondary reviewers were used to check the accuracy of screening, data extraction and risk-of-bias assessment processes. Due to the scale of the project, this systematic review was performed in conjunction with a team of secondary reviewers and an information specialist (see Acknowledgments and Author declaration). <!-- TODO add the following to acknowledgements These included Alexandra MacAleenan, Athena Sheppard, and Matthew Lee.  In addition, Sarah Dawson, an information specialist, provided input to the design of the search strategy.-->

@@

### Search strategy

I systematically searched several electronic bibliographic databases to identify potentially relevant entries (hereafter referred to as "records"). The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. As the contents of the Web of Science Core Collection can vary by institution,[@gusenbauer2020a] the specific databases searched via this platform are listed in Appendix \@ref(appendix-wos-databases). The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy included terms related to lipids, lipid modifying treatments, and dementia, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

I also searched clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials. In addition, I searched the bioRxiv and medRxiv preprint repositories using the tool developed in Chapter \@ref(sys-rev-tools-heading) to identify potentially relevant preprinted studies.

Grey literature was searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses were accessed using the Open Access Theses and Dissertations portal. In addition, the abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) were searched by hand. <!-- TODO how were these searched? Date and other range --> Finally, the reference lists of included studies were searched by hand while studies citing included studies was examined using Google Scholar (forward and reverse citation searching or "snowballing"<!-- TODO CITATION NEEDED -->).

### Study selection

Records were imported into Endnote and de-duplicated using the method outlined in Bramer et al. (2016).[@bramer2016] In summary, this method uses multiple stages to identify potential duplicates, beginning with automatic deletion of records matching on multiple fields ("Author" + "Year" + "Title" + "Journal"), followed by manual review of less similar articles (e.g. those matched based on the "Title" field alone).

Following deduplication of records, screening (both title/abstract and full-text) was performed using a combination of Endnote, a citation management tool,[@hupe2019] and Rayyan, a web-based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records was performed primarily by me, with a random ~10% sample of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. Additionally, records were rescreened by me with a 1 month lag to intra-rater consistency.

Similarly, I completed all full-text screening, with a random ~10% being screened in duplicate by a second reviewer. In addition, any records identified I identified as being difficult to assess against the inclusion criteria were screened in duplicate. Reasons for exclusion at this stage were recorded. Disagreements occurring during either stage of the screening process were resolved through discussion with a senior colleague. A PRIMSA flow diagram was produced to document how records moved through the review.[@page2021]

The criteria used to assess eligibility are presented in the subsequent sections.

@@

I sought to include studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia and its subtypes. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants were screened for dementia at baseline and prevalent cases excluded. Alternatively, where no baseline screening was employed, participants were assumed to be dementia free if less than <50 years of age at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits were placed on the sample size of included studies.

Eligible studies defined dementia according to recognised criteria, for example the International Classification of Diseases (ICD),[@organizationwho1993] National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN),[@roman1993] or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria.[@edition2013] Studies utilising electronic health records were the exception to this, as it was assumed that these criteria were used when entering the outcome into the EHR.

<!-- TODO Cite Peter Tennant's piece here on the problems with causal inference from analysis of change scores -->

<!-- TODO Any study using EHR - go back and extract codelists -->

No limitations were imposed on publication status, publication date, venue or language, although sufficiently detailed reports of the studies to be able to examine their methods were required for inclusion.

@@

Due to the significant impact of a memory-related outcome such as dementia on exposure recall, case-control studies were excluded, though nested case-control studies, where historical records are used to determine the exposure status, were eligible for inclusion. Cross-sectional studies, qualitative studies, case reports/series and narrative reviews were also excluded. Studies which presented no evidence of attempting to exclude prevalent cases from their analyses were also excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/dementia) were excluded. Conference abstracts with no corresponding full-text publication were examined, and where required, I contacted authors to obtain information on the study's status. Previous systematic reviews were not eligible for inclusion, but their reference lists were screened to identify any potentially relevant articles.

Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a "multi-domain intervention" where a lipid-regulating agent is included in each arms (e.g. for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e. below the age of 50) were excluded. Finally, studies using a dietary intervention, for example omega-3 fatty acid enriched diet, were excluded as it is difficult to disentangle the effect of other elements contained within the diet. Note, this is distinct from studies which delivered a simple tablet-based omega-3 intervention, which would have been eligible for inclusion.

### Validation of screening process

Inter- and intra-rater reliability during the screening stages were assessed for a 10% sub-sample of records. Intra-rater reliability involved a single reviewer applying the inclusion criteria to the same set of records while blinded to their previous decisions (i.e. assessment of conistency), while inter-rater reliability involved two reviewers independently screening the same set of records (i.e. assessment of accuracy).

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is an important characteristic.

@@

Intra- and inter-rater reliability was assessed against these cut-offs. If this assessment demonstrated issues with the screening process (defined as an AC1 of <.9), a larger proportion of records would have been dual-screened.

@@

Data extraction was performed using a piloted data extraction form. Extracted items included: article metadata (year of publication, author list, journal), study characteristics (study location, data source, exposure, outcomes, outcome criteria used), patient characteristics (age, sex, baseline cognition scores, baseline education scores), and results (exposure-outcome pairing, effect measure, effect estimate, error estimate, p-value). I extracted all data in the first instance, which was subsequently checked for accuracy by a second member of the review team.

#### Study-fication

As part of the data extraction process, multiple resords resulting from the analysis of the same data were included and grouped into single units, hereafter called studies. This is likely in the advent of multiple papers reporting results on the same cohort, but say, at different time points. Study-fication builds out the most comprehensive accounts of the studies and results from as many published articles were applicable.

This was particularly relevant to preprints and published papers reporting the same study, which were not considered to be duplicate records but instead different reports of the same study. This is due to the potential for the published version to offer some information that the preprint did not, and vice versa.

#### Combining across groups

Following best practice, where summary data was presented across two groups (e.g. age at baseline stratified by hypercholesterolemia status), the following approach was used to combine the groups. [@higgins2019]

\begin{equation}
N = N_1 + N_2
  (\#eq:combiningGroups1)
\end{equation}

\begin{equation}
Mean = \frac{(N_1M_1 + N_2N_2)}{(N_1 + N_2)}
  (\#eq:combiningGroups2)
\end{equation}

\begin{equation}
SD = \sqrt{\frac{(N_1-1)SD_1^2 + (N_2-1)SD_2^2 + \frac{N_1N_2}{N_1 + N_2}(M_1^2 + M_2^2 - 2M_1M_2)}{N1 + N2 -1}}
  (\#eq:combiningGroups3)
\end{equation}

This was implemented in a systematic manner, with the raw group data being extracted and a cleaning script written to combine the groups for analysis.

#### Harmonisation of cholesterol measures

<!-- TODO This might not be as accurate now given the focus on dose response --> <!-- TODO Cross reference with the data extraction issues here - i.e. not providing qratile cut-off values, not poviding numbers per group, etc. --> Where necessary, lipid levels reported in _mmol/L_ were converted in _mg/dL_ using the following formula:

@@

@@

Where additional data points not included in the report of an analysis were required either for the analysis or risk-of-bias assessment, the corresponding author of the study was contacted. This approach was taken due to the potentially large impact of following up with authors on the results of the review.[@reynders2019]

#### Analysis of varying effect measures

<!-- TODO Include formulae and informed assumptions  -->

The range of effect measures presented by studies (odds ratios, risk ratios, hazard ratios, etc) are not directly interchangeable in the context of systematic review. If the outcome is rare, as is the case for dementia outcomes at and estimated prevalence of  odds and risk ratios approximate each other.<!-- TODO CITATION NEEDED --> However, hazard ratios provide a very different interpretation, taking into account person-timea-at-risk in each treatment group. As such, different effect estimates can be one potential problem that precludes a meta-analysis of all studies.[@mckenzie2019]

Several existing reviews do not distinguish between the types of effect measures and include all existing studies in a single meta-analysis to produce an overall effect. In addition, there is some evidence of manipulation of effect estimates in previous reviews,(e.g. Chou, Sci Reports - at least one study disagrees with) but this is not accurately documented in the review text.
<!-- TODO Need to really fix this -->

In this review, studies reporting hazard ratios were synthesised separately to those reporting odds/risk ratios.

### Risk-of-bias assessment {#risk-of-bias}

A key use of the review presented in this chapter is to identify different sources of evidence at risk of a diverse range of biases, and to contrast and compare findings across them (see Section \@ref(triangulation-overview) for an overview of triangulation and Section \@ref(intro-triangulation) for the results of this qualitative analysis). To enable this triangulation exercise, a detailed and structured risk-of-bias assessment formed an important part of this review.

@@

This move was prompted by a unclear definition of "methodological quality" which could include facets such as unclear reporting, and challenges in the comparison of results from different tools. <!-- TODO CITATION NEEDED --> As part of this shift, the community also moved from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order. Additionally, bias should be assessed at the result (defined as a a specific outcome at a specific timepoint) rather than the study level. For example, a study may report on the efficacy of an intervention at six months and two years follow-up. In this case, missing outcome data that is not an issue at six months may introduce bias at 2 year follow-up, and assigning a bias judgement to the study as a whole masks the different level of bias for each unique result.

@@

@@

An inital qualiatative synthesis of evidence was performed, summarising the data extracted from studies stratified by study design

Where individual studies were deemed comparable, they were incorporated into a quantitative analysis or "meta-analysis". <!-- TODO CITATION NEEDED --> Meta-analysis provides a summary or pooled effect estimate across studies.
